---
title: "p8105_hw3"
author: "Qianying Wu"
date: "2023-10-12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(p8105.datasets)
```

## Problem 1

1. Read in the data

```{r}

data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

2. Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row  a single product from an instacart order. Variables are identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

### Data cleaning

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}

library(p8105.datasets)
data("brfss_smart2010")
brfss = brfss_smart2010 |> 
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent")) |> 
  rename(state = locationabbr, county = locationdesc)

brfss
#brfss = brfss_ori |> filter(response == "Excellent" | response == "Poor")
#head(brfss_ori)
```


1. In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
state_2002 <- brfss |> filter(year == "2002") 


state_2002_7 = state_2002 |> 
  group_by(state) |>
  summarise(occurences = n_distinct(county)) |>
  filter(occurences >= 7)

state_2002_7

state_2010 <- brfss |> filter(year == "2010") 


state_2010_7 = state_2010 |> 
  group_by(state) |>
  summarise(occurences = n_distinct(county)) |>
  filter(occurences >= 7)

state_2010_7
```

- According to the result, we know that in 2002, the states CT, FL, MA, NC, NJ, PA were observed at 7 or more locations. 

- In 2010, the states CA, CO, FL, MA, MD, NC, NE, NJ, NY, PA, SC, TX, WA were observed at 7 or more locations. 


2. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}

excellent_data <- brfss |>
  filter(response == "Excellent") |>
  select(year, state, data_value)

# Calculate the average data value within each state for each year
average_data_by_state <- excellent_data |>
  group_by(year, state) |>
  summarise(average = mean(data_value))

average_data_by_state
# Create a "spaghetti" plot
ggplot(average_data_by_state, aes(x = year, y = average, group = state, color = state)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Average Data Value",
    title = "Average Data Value Accross Year In Different States"
  )
```

* Comment: From the plot, we can see the average data value across year in different states. For all states, we generally see a low average data value at the year of 2005. 

3. Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.


```{r}
brfss |>
  filter(year %in% c(2006, 2010), 
         state == "NY", 
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |>
         ggplot(aes(x = response, y = data_value)) +
         geom_boxplot() +
         facet_wrap(~year, ncol = 2) +
          labs(
        x = "Data Value",
        y = "Frequency",
        title = "Distribution of Data Value by Response in NY State (2006 and 2010)",
        fill = "Response"
  )

```
* Comment: From the boxplot, for both years, we can see that in general, the frequency of poor response is the lowest response, and the most frequent response for the data_value would be "Very good" in both years. The "good" response is the second most frequent response. 

## Problem 3

1. Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
accel = read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  pivot_longer(min1:min1440,
               names_prefix = "min",
               names_to = "minutes",
               values_to = "MIMS")

accel
  

demo = read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  drop_na() |>
  filter(age >= 21) |>
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "less than high school", "2" = "high school", "3" = "more than high school"))

redemo = demo |>
    mutate(sex = fct_relevel(sex, "male", "female"),
           education = fct_relevel(education, "less than high school", "high school", "more than high school"))

redemo

nhanes = left_join(redemo,accel,by = "seqn")
```

2. Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
nhanes |> 
  group_by(sex, education) |>
  summarise(total_num = n()/1440) |>
  pivot_wider(
    names_from = sex,
    values_from = total_num
  )  |>
  knitr::kable(caption = "Summary of Participants by Sex and Education")

# Visual distribution

age_distribution_plot <- nhanes |>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(x = "Education", y = "Age") +
  scale_fill_manual(values = c("male" = "purple", "female" = "green")) +
  facet_wrap(~sex)

print(age_distribution_plot)

```

* Comment: 

On this visualization, it is clear that in the education level of less than high school, both female and male groups have similar distribution including their mean age at around 60. The female group has their mean age slightly larger than the male's group.

In equivalent to high school education, female group has mean age larger than 60, which is higher than the mean age of the male group's distribution (slightly above 50).

In the more than high school education level, both female and male groups have similar distributions of age, with male group's mean age (around 42) slightly higher than those of the female group (around 41). 

From the male groups, we can see that when the education level becomes higher, the mean age of the male's group becomes younger. From the female groups, we observe that for lower than high school and equivalent to high school groups, the mean age are similar, which is much larger than the mean age of higher than high school education level. 


3. Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.


```{r}

nhanes |>
  group_by(seqn, education, sex, age) |>
  summarise(n = sum(MIMS)) |>
  ggplot(aes(x = age, y = n, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education) +
  labs(x = "Age", y = "Total Activity") +
  scale_color_manual(values = c("male" = "purple", "female" = "green"))


```

* comment: 
From the plot, we can see that the overall total activity in three different education levels all show a decreasing trend when age increased in both male and female groups. 

In less than high school education group, female group has a higher total activity when age is less than 40. However, between age 40 to age 80, the male group has higher total activity. For both sex, they have a peak at around age 60.

In the equivalent to high school education, the trend of the female group has a higher total activity than the male group. For both sex groups, they have their peak at age 40.

In the more than high school education group, he trend of the female group has a higher total activity than the male group. 

4. Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
nhanes |>
  mutate(minutes = as.numeric(minutes)) |>
  ggplot(aes(x = minutes, y = MIMS, color = sex)) +
  geom_line(alpha = 0.2) +
  geom_smooth(aes(group = sex), se = FALSE) +
  facet_wrap(~ education) +
  labs(x = "Minutes", y = "MIMS (Activity)") +
  scale_color_manual(values = c("male" = "purple", "female" = "green"))


```

* Patterns: 

From the graph, it is clear that both age groups have the similar trend of MIMS activity across the time of day in less than education and equivalent to education groups, in which they have their minimum value in the early  time of the day. In the more than education group, although male and female group also share the similar trend as the other two education groups with their minimum value of MIMS activities in the middle time of the day, there are some slight difference between the two gender groups. The female group has higher MIMS activities in the early to middle times of the day, and the male group has higher MIMS activities in the late times of the day. 